\chapter{Outline}

https://swathisprasad.medium.com/beginners-guide-to-java-virtual-threads-with-spring-boot-753bd84a643d

Chapter 1: File Operations and Java Collections Framework
Basics of File Handling: java.io.File, reading and writing files.
Advanced File Operations: java.nio, Path, Files.
Java Collections Framework: Overview, Lists, Sets, Maps, and their implementations.
Generic Classes and Methods: Creating and using generic types.
Exercises: File operations, implementing generic methods, and working with different types of collections.
Chapter 2: Advanced Streams, Lambda Expressions, and Multithreading
Streams and Lambda Expressions: Introduction, advanced operations.
Multithreading Basics: Creating threads, Runnable, Callable, Future.
Thread Synchronization: synchronized, Locks, thread-safe collections.
Parallel Streams: Using streams in a multithreaded context.
Exercises: Lambda expressions, stream manipulations, basic multithreading tasks.
Chapter 3: Build Automation with Maven
Maven Basics: Project Object Model (POM), dependencies, plugins.
Maven Lifecycle: Build phases, goals.
Creating and Running Java Applications: Packaging, running tests.
Exercises: Setting up a Maven project, adding dependencies, customizing the build process.
Chapter 4: Unit Testing with JUnit, Mockito, and MockMvc
JUnit: Writing and running tests, assertions, test lifecycle.
Mockito: Creating mocks, stubbing methods, verifying interactions.
MockMvc: Testing Spring MVC applications, simulating HTTP requests.
Exercises: Writing unit tests for different layers (service, repository, controller) of an application.
Chapter 5: JDBC, JPA, and Spring Data JPA
From JDBC to JPA: Basics of JDBC, introduction to JPA.
Spring Data JPA: Repositories, query methods, custom queries.
Entity Relationships: Mapping associations in JPA.
Exercises: Implementing CRUD operations using JDBC, JPA, and Spring Data JPA.
Chapter 6: Servlet Technology and Spring Security Basics
Servlet Basics: Servlet lifecycle, HTTP servlets, request-response model.
Spring Security Fundamentals: Authentication, authorization, configuring security.
Exercises: Creating a basic servlet, securing a web application with Spring Security.
Chapter 7: Spring Boot and Advanced Topics
Spring Boot Essentials: Auto-configuration, starters, properties.
File Upload and PDF Generation: Using Spring Boot for file operations.
Asynchronous Operations: @Async, handling asynchronous processes.
Relevant Topics from Blogs and Java Magazine: Latest trends, best practices, new features in Java and Spring Boot.
Final Project: Building a full-stack application incorporating the covered topics.
Additional Exercises and Projects:
Build a Blogging Platform: Incorporate JPA for data handling, Spring Security for user authentication, and Spring MVC for the web interface.
E-commerce Backend: Use Spring Boot, JPA, and Spring Security to simulate an e-commerce backend with user authentication and product management.
Multi-threaded File Processing Application: Create an application that processes large files in a multithreaded manner using Java's concurrency tools.